Initialization for missing data

### 2017-09-16 analysis/index.Rmd
Files commited by wflow_commit.

Click on this [summary of the rank0 vs rank1](Obj_summary.html) to see more details.

@@

### meeting summary

[04-26](meetnotes.html)

>>>>

Click on this [summary of the rank0 vs rank1](Obj_summary.html) to see a simple summary to the following two items.

@@

## flashr2

This is a simple test [first try](flash_beta_test.html)
Initialization issue: [softimpute as initialization](missingimputeinitial.html)

### 2017-09-16 analysis/missingimputeinitial.Rmd
Files commited by wflow_commit.

---
title: "test flashr2"
author: "Wei"
date: 09-12
output: 
  html_document:
    code_folding: hide
---

@@

### this is code for simulated data

@@

## check the prediction performance, this is something I don't understand

In this case I run flash for different situation and I found the old version has smaller RMSE.

I print the obj value to check if it is increase.

```{r}
library(PMA)
library(softImpute)
set.seed(999)
N = 100
P = 200
dat = sim_K(K=1,N, P, SF = 0.5, SL = 0.5, signal = 1,noise = 5)
# a = rchisq(N,3)
# b = rchisq(P,1)
# data = sim_hd(N, P, SF = 0.5, SL = 0.5, signal = 1, a ,b,mu = 0)
Y = dat$Y
Y_miss = Y
Y_miss[1:50,1:100] = NA
Y_miss[51:100,101:200] = NA
```

```{r}
data = set_flash_data(Y)
g1= flash_r1(data,verbose=TRUE)
gg1 = flashr::flash(Y)
sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(gg1$l %*% t(gg1$f)))^2)
sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(g1$EL %*% t(g1$EF)))^2)
sum((as.vector(dat$L_true %*% t(dat$F_true))-0)^2)
```

### for this case, when data are missing, the obj is not increase

the final result is rank zero, so I am not sure it is a problem

```{r}
data.miss = set_flash_data(Y_miss)
g1.miss = flash_r1(data.miss,verbose=TRUE)
```

## missing value

In this case, I have two problems:

1. The Obj function seems not increasing.

2. the RMSE seems a bit large for flashr2. I think the $\tau$ estimation is not that precise.

```{r}
library(PMA)
library(softImpute)
set.seed(999)
N = 100
P = 200
dat = sim_K(K=1,N, P, SF = 0.5, SL = 0.5, signal = 3,noise = 5)
# a = rchisq(N,3)
# b = rchisq(P,1)
# data = sim_hd(N, P, SF = 0.5, SL = 0.5, signal = 1, a ,b,mu = 0)
Y = dat$Y
Y_miss = Y
Y_miss[1:50,1:100] = NA
Y_miss[51:100,101:200] = NA
```

```{r}
data.miss = set_flash_data(Y_miss)
g1.miss = flash_r1(data.miss,verbose=TRUE)
gg1.miss = flashr::flash(Y_miss)
sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(gg1.miss$l %*% t(gg1.miss$f)))^2)
sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(g1.miss$EL %*% t(g1.miss$EF)))^2)
sum((as.vector(dat$L_true %*% t(dat$F_true))-0)^2)
```

>>>>

---
title: "Initialization for missing data"
author: "Wei Wang"
date: 09-16
output: 
  html_document:
    code_folding: hide
---

@@

@@

## same data set as in "flash_beta_test"

```{r,message=F,warning=F}
library(softImpute)
set.seed(999)
N = 100
P = 200
dat = sim_K(K=1,N, P, SF = 0.5, SL = 0.5, signal = 3,noise = 5)
# a = rchisq(N,3)
# b = rchisq(P,1)
# data = sim_hd(N, P, SF = 0.5, SL = 0.5, signal = 1, a ,b,mu = 0)
Y = dat$Y
Y_miss = Y
Y_miss[1:50,1:100] = NA
Y_miss[51:100,101:200] = NA
```

Here I just run flash on this data sets four times and provide different result due to the unstable initial value.

```{r}
par(mfrow = c(2,2))
for(i in 1:4){
  data.miss = set_flash_data(Y_miss)
  g1.miss = flash_r1(data.miss,verbose=F)
  plot(as.vector(dat$L_true %*% t(dat$F_true)), as.vector(g1.miss$EL %*% t(g1.miss$EF)), xlab = "truth",ylab = "estimated", main = "vectorized LF")
}
```

## I don't think softimpute is good initialization method (or maybe I don't understand this method very well)

Here is a example of use softimpute. maybe I was not using it in the right way.

Here I use default `lambda`, and we can try `type = "svd"`. I tried different combination and this method is still not stable.

```{r}
par(mfrow = c(2,2))
for(i in 1:4){
  gsoft = softImpute(Y_miss, rank.max = 1,type = "als")
  sfl = (gsoft$u)
  sff = gsoft$d * (gsoft$v)
  plot(as.vector(dat$L_true %*% t(dat$F_true)), as.vector(sfl %*% t(sff)), xlab = "truth",ylab = "estimated", main = "vectorized LF")
}
```
